# Explore - Scraping

## Goals

Given a Billboard chart slug name (from it's URL) and date (default to current), scrape the chart and save the file.

Charts I'm using this on are [Billboard Hot 100](https://www.billboard.com/charts/hot-100/) and [Billboard 200](https://www.billboard.com/charts/billboard-200/).

A compilation of the resulting files are handled in a different script.

> This same concept is used with the Github Action version that scrapes only the current chart: `action_scrape_charts.R`.

> The "write_rds" step is commented out so files are not overwritten unless purposely changes.

## Setup

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(rvest)
library(lubridate)
library(here)
```


## Set up the list of charts

This comes from the url of the chart, like "hot-100" from `https://www.billboard.com/charts/hot-100/`.

```{r charts-list}
charts_list <- c("hot-100", "billboard-200")
```

## Date options

I set up some flags to get a specific date or the current date. I won't need this with the final Github Action

- `F` pulls the current chart
- `T` pulls the chart that is noted in the `edition_request` object

```{r flags}
edition_request <- "2026-09-19"
edition_flag <- F
```


## Scrape a single file

```{r}
main_url <- "https://www.billboard.com/charts/hot-100/"

# get initial scrape
scrape_first <- read_html(main_url)

# pull date from scrape
chart_date <- scrape_first |> 
  html_element("div#chart-date-picker") |> 
  html_attr("data-date")

# pull year from scrape
chart_year <- year(chart_date)

# process results into a tibble
scrape_tibble <- scrape_first |> 
  html_elements("ul.o-chart-results-list-row") |> 
  html_text2() |> 
  as_tibble()
```

Each line has some extra columns repeated ... LW, PEAK, WEEKS ON CHART. On the first line for the number 1 song, there is also WEEKS AT NO. 1. This was added the chart week of Sept. 19th when the top song tied for the longest. I'm not sure if it will remain a part of the chart, but I'd like to deal with it.

### Process first row

Here we grab the first row, then check if the term "NO. 1" exists, and if so process it as a 7-variable table. If not, then we process as a 6-variable table.

```{r}
scrape_toprow <- scrape_tibble |>
  slice(1) |> 
  mutate(
    data_cleaned = str_remove_all(value, " NEW\nNEW|LW\n|PEAK\n|WEEKS ON CHART\n|WEEKS AT NO. 1\n|WEEKS\n"),
    data_cleaned = str_remove_all(data_cleaned, " RE- ENTRY\n|RE- ENTRY|RE-ENTRY"),
    data_cleaned = str_extract(data_cleaned, "^((?:[^\n]*\n){6}[^\n]*)") # sevens
  ) |> 
  select(data_cleaned, value) |>
  separate(
    col = data_cleaned,
    sep = "\n",
    into = c(
      "current_week",
      "title",
      "performer",
      "last_week",
      "peak_pos",
      "wks_at_no1",
      "wks_on_chart"
    )
  ) |>
  select(-value) |>
  mutate(chart_week = chart_date, .before = current_week)
```

Writing

### Process other rows

```{r}
scrape_rest <- scrape_tibble |>
  slice(2:100) |> 
  mutate(
    data_cleaned = str_remove_all(value, " NEW\nNEW|LW\n|PEAK\n|WEEKS ON CHART\n|WEEKS AT NO. 1\n|WEEKS\n"),
    data_cleaned = str_remove_all(data_cleaned, " RE- ENTRY\n|RE- ENTRY|RE-ENTRY"),
    data_cleaned = str_extract(data_cleaned, "^((?:[^\n]*\n){5}[^\n]*)") # sevens
  ) |> 
  select(data_cleaned, value) |>
  separate(
    col = data_cleaned,
    sep = "\n",
    into = c(
      "current_week",
      "title",
      "performer",
      "last_week",
      "peak_pos",
      # "wks_at_no1",
      "wks_on_chart"
    )
  ) |>
  select(-value) |>
  mutate(chart_week = chart_date, .before = current_week)
```

### Bind them together

```{r}
scrape_clean <- scrape_toprow |> 
  bind_rows(scrape_rest)

scrape_clean
```


## Scraper function starts here

```{r}
for (chart in charts_list) {

  # Sets edition flag
  edition_current <- ""
  if (edition_flag == T) edition_page <- edition_request else edition_page <- edition_current

  # make the url
  scrape_url <- paste(
  "https://www.billboard.com/charts/",
  chart,
  "/",
  edition_page,
  sep = ""
  )
  
  # get initial scrape
  scrape_first <- read_html(scrape_url)
  
  # pull date from scrape
  chart_date <- scrape_first |> 
    html_element("div#chart-date-picker") |> 
    html_attr("data-date")
  
  # pull year from scrape
  chart_year <- year(chart_date)
  
  # process results into a tibble
  scrape_tibble <- scrape_first |> 
    html_elements("ul.o-chart-results-list-row") |> 
    html_text2() |> 
    as_tibble()

  # process first row
  scrape_toprow <- scrape_tibble |>
    slice(1) |> 
    mutate(
      data_cleaned = str_remove_all(value, " NEW\nNEW|LW\n|PEAK\n|WEEKS ON CHART\n|WEEKS AT NO. 1\n|WEEKS\n"),
      data_cleaned = str_remove_all(data_cleaned, " RE- ENTRY\n|RE- ENTRY|RE-ENTRY"),
      data_cleaned = str_extract(data_cleaned, "^((?:[^\n]*\n){6}[^\n]*)") # sevens
    ) |> 
    select(data_cleaned, value) |>
    separate(
      col = data_cleaned,
      sep = "\n",
      into = c(
        "current_week",
        "title",
        "performer",
        "last_week",
        "peak_pos",
        "wks_at_no1",
        "wks_on_chart"
      )
    ) |>
    select(-value) |>
    mutate(chart_week = chart_date, .before = current_week)
  
  # process rest of rows
  scrape_rest <- scrape_tibble |>
    slice(2:100) |> 
    mutate(
      data_cleaned = str_remove_all(value, " NEW\nNEW|LW\n|PEAK\n|WEEKS ON CHART\n|WEEKS AT NO. 1\n|WEEKS\n"),
      data_cleaned = str_remove_all(data_cleaned, " RE- ENTRY\n|RE- ENTRY|RE-ENTRY"),
      data_cleaned = str_extract(data_cleaned, "^((?:[^\n]*\n){5}[^\n]*)") # sevens
    ) |> 
    select(data_cleaned, value) |>
    separate(
      col = data_cleaned,
      sep = "\n",
      into = c(
        "current_week",
        "title",
        "performer",
        "last_week",
        "peak_pos",
        # "wks_at_no1",
        "wks_on_chart"
      )
    ) |>
    select(-value) |>
    mutate(chart_week = chart_date, .before = current_week)
  
  # bind them
  scrape_clean <- scrape_toprow |> 
    bind_rows(scrape_rest)

  # name path to save file
  folder_path <- paste("data-scraped/", chart, "/", chart_year, "/", sep = "")
  
  # create the directory if it doesn't exist
  if (!dir.exists(here(folder_path))) {dir.create(here(folder_path), recursive = TRUE)}
  
  # write the file
  scrape_clean |> write_csv(paste(folder_path, chart_date, ".csv", sep = ""))

}
```
