---
title: "Combine for current charts"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
---

## Goals of this notebook

To combine scraped Billboard chart files with a previous-processed archive to create a "current" archive. See `01-scrape-charts.Rmd` for more info on scraping, and the README for more about the archives.

## Setup

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(fs)
library(lubridate)
```

## Chart list

Based on url slug.

```{r charts-list}
charts_list <- c("hot-100", "billboard-200")
```

## One-time processing

This combination script does require some **one-time** processing of some previously-worked files (see README).

The code is **commented** because it doesn't need to run again.

```{r archive-process}
# archive_process_hot100 <- read_csv("data-out/hot100_archive_1958_2021.csv") |> 
#   rename(
#     chart_week = chart_date,
#     current_week = current_position,
#     last_week = previous_position,
#     peak_pos = peak_position,
#     wks_on_chart = weeks_on_chart
#   )
# 
# archive_process_bb200 <- read_csv("data-out/billboard200.csv") |> 
#     select(
#       chart_week = date,
#       current_week = current,
#       title,
#       performer = artist,
#       last_week = previous,
#       peak_pos = peak,
#       wks_on_chart = weeks
#     )
# 
# archive_process_hot100 |> write_csv("data-scraped/hot-100/previous_archive.csv")
# archive_process_bb200 |> write_csv("data-scraped/billboard-200/previous_archive.csv")
```


## The loop

Here we combine all the files for each chart and (in the darkness) bind them.

Of the many ways to do this, this is the sparsest binding code. Not sure what I would do it this way if read_csv needed arguments. This uses the tidyverse purrr [map](https://purrr.tidyverse.org/reference/map.html) function.


```{r combine-loop}

for (chart in charts_list) {

  # compile a list of files to scrape
  files_list <-  dir_ls(paste("data-scraped", chart, sep = "/"), recurse = TRUE, regexp = ".csv")

  # combine all the files
  charts_recent <- map_dfr(files_list, read_csv, col_types = cols("last_week" = col_integer()))
  
  # write the combined file
  charts_recent |> write_csv(paste("data-out/", chart, "-current.csv", sep = ""))
  
}

```

