---
title: "RWDIR Hot 100 data"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

## Goal of this notebook

The goal here is to create Billboard Hot 100 data for the [Reporting with Data in R book](https://utdata.github.io/rwdir/). It is purposely mucked up for the assignment.

> In June 2023 the script `action_combine_charts.R` was updated to create assignment data from "current" data. So this isn't really needed anymore except for posterity.

## Setup

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(lubridate)
```

## Import data

Import the most current Billboard Hot 100 data.

```{r import}
hot100_current <- read_csv("data-out/hot-100-current.csv")
```

## Clip the year

For the book assignment, we want only full years so we remove rows from the most recent year.

> This is bad ... we don't clip the year anymore

```{r clip_year}
# This gets the current year.
hot100_clipped <- hot100_current |> 
  filter(year(chart_week) < year(Sys.Date()))

hot100_clipped$chart_week |> summary()
```


## Muck the data

We change the names of the column headers and muck up the data to force students to deal with that in [Reporting with Data in R Chapter 3](https://utdata.github.io/rwdir/counts-import.html#cleaning-data).


```{r muck}
hot100_assignment <- hot100_clipped %>%
  # muck the date
  mutate(
    chart_week = paste(
      month(chart_week) %>% as.character(),
      day(chart_week) %>% as.character(),
      year(chart_week) %>% as.character(),
      sep = "/"
      )
  ) |>   
  # muck the names
  select(
    `CHART WEEK` = chart_week,
    `THIS WEEK` = current_week,
    `TITLE` = title,
    `PERFORMER` = performer,
    `LAST WEEK` = last_week,
    `PEAK POS.` = peak_pos,
    `WKS ON CHART` = wks_on_chart
  )

hot100_assignment %>% glimpse()
```

## Export

Here we export the data for the assignment.

This is commented out so I don't overwrite "current" data.

```{r export}
# hot100_assignment |> write_csv("data-out/hot100_assignment.csv")
```

