---
title: "Various data checks"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---


## Goals

Check for problems in the combined current archives

## Setup

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Hot 100

Import

```{r}
hot100 <- read_csv("data-out/hot-100-current.csv")

hot100 %>%  glimpse()
```

Check most recent chart

```{r check}
counts_hot100 <- hot100 %>% 
  group_by(chart_week) %>% 
  summarise(cnt = n()) %>% 
  arrange(chart_week %>% desc())

counts_hot100 %>% head()
```

Look for short counts

```{r short-counts}
counts_hot100 %>% 
  filter(cnt != 100)
```

## Billboard 200

Import

```{r bb200-import}
bb200 <- read_csv("data-out/billboard-200-current.csv")

bb200 |> glimpse()
```

See most recent charts

```{r}
bb200 |> 
  count(chart_week) |> 
  arrange(chart_week |> desc()) |> 
  head()
```

Check counts less than 200

```{r}
bb200 |> 
  count(chart_week) |> 
  filter(n != 200)
```


- The 175 weeks are how they appear online today. Could be that was all there were as that was the chart's birth.
- The `1967-09-16` data is an error in my collected data that needs to be fixed, but I haven't decided how yet.
