---
title: "Billboard Hot 100 data processing"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
---

This notebook downloads Billboard Hot 100 data from data.world and simplifies it for later assignments. There is a python notebook attempting to pull this data from the original source, but there were problems and this route seemed easier.

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
library(lubridate)
```


## Data from data.world

Weekly Hot 100 singles chart between 8/2/1958 through 2020 as pulled from [data.world](https://data.world/kcmillersean/billboard-hot-100-1958-2017). See link for more information.

### Download and store

This is commented once done.

```{r download}
df <- read.csv("https://query.data.world/s/zshes27odrcz75zgjts35bpyfwurnf", header=TRUE, stringsAsFactors=FALSE);

df %>% write_rds("../data-process/hot-100/hot100-orig.rds")
df %>% write_csv("../data-process/hot-100/hot100-orig.csv")
```

### Import

```{r import}
hot100 <- read_rds("../data-process/hot-100/hot100-orig.rds") %>% clean_names()

hot100 %>% glimpse()
```

## Create real date

```{r}
hot100_date <- hot100 %>% 
  mutate(
    week_id = mdy(week_id)
  ) %>%
  arrange(week_id, week_position)

hot100_date %>% glimpse()
```

## Create a clean, smaller version

```{r}
hot100_tight <- hot100_date %>%
  rename(
    date = week_id,
    title = song,
    artist = performer,
    current = week_position,
    previous = previous_week_position,
    peak = peak_position,
    weeks = weeks_on_chart
  ) %>% 
  select(date, current, title, artist, previous, peak, weeks) %>% 
  arrange(date, current)

hot100_tight %>% glimpse()
```

## Summary

```{r}
hot100_tight %>% summary()
```


## Export the cleaned data

```{r}
hot100_date %>% write_rds("../data-out/bb_hot100.rds")
hot100_date %>% write_csv("../data-out/bb_hot100.csv")
hot100_tight %>% write_rds("../data-out/bb_hot100_tight.rds")
hot100_tight %>% write_csv("../data-out/bb_hot100_tight.csv")
```

